---
title: "Solar Detection & Income"
format: html
editor: visual
---

### Libraries

```{r}
library(tidyverse)
library(here)
library(sf)
library(mapview)
install.packages("tidycensus")
library(tidycensus)
```

### Read in Data

```{r}
oahu_solar <- st_read(here("data/oahu_solar_dedup_simple.geojson"))
```

### Map the data

```{r}
mapview(oahu_solar)
```

### Retreive some census data

```{r}
oahu_median_household_income <- get_acs(geography = "tract",
                                 variable = "B19013_001",
                                 state = "HI",
                                 county = "Honolulu",
                                 geometry = TRUE,
                                 year = 2022)
```

```{r}
mapview(oahu_median_household_income,
        zcol = "estimate")
```

```{r}
ggplot(oahu_median_household_income, aes
       (x = moe, y = estimate)) +
  geom_point() +
  geom_smooth(method = "lm")
```

- package: sf / language: R

- head() of two databases

- please identify the best function in sf for the task at hand

- Goal is to answer the question: "Is there a relationship/correlation between median income and solar infrastructure at the census tract level"?

- one census tract level

- i got out median income data from tidycensus

- building

```{r}
oahu_solar_clean <- oahu_solar |>
  filter(!st_is_empty(geometry)) |>
  st_make_valid() |>
  st_transform(crs = 26904)
# drop the 68 empty geometries
# repair any invalid polygons
# NAD83 / UTM zone 4N (meters)

# Project the census tracts to the same CRS
oahu_tracts <- oahu_median_household_income |>
  st_transform(crs = 26904)

# Confirm both layers now share a CRS (should print TRUE)
st_crs(oahu_solar_clean) == st_crs(oahu_tracts)
```

```{r}

# Represent each solar polygon with one point guaranteed to lie inside it
oahu_solar_points ‹- oahu_solar_clean |>
  st_point_on_surface ()
# Attach each point to the census tract it falls in
solar_with_tracts <-oahu_solar points |> st_join(
  oahu_tracts> select (GEOID),
  join = st
  _intersects
  #bring over only the tract ID
  # match points to the tract they touch
  # Sanity checks
  nrow(solar_with_tracts)
  sum(is.na(solar with_tracts$GEOID))
```
